Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-09 Chapter-1/1-Theory.tex
s

\subsection{Why so sneaky? Neutrino Cross Section Challenge}

@@

\textbf{Majorana or Dirac?}
Evidence of neutrino oscillations demand the introduction in the theory of a mechanism which can give mass to the neutrinos.  This mechanism should possibly also explain why neutrino masses are at least six orders of magnitude lower than the electron mass (the second lightest SM fermion).
In a description of neutrinos as Dirac 4-component spinors, the neutrino field acquires mass  via the Higgs mechanism as any other fermion of the SM. In this case, the neutrino mass is given by 
%\begin{equation} 
$m_a = \frac{y^\nu_a v}{\sqrt2}$, %\end{equation} 
where $v$ is the Higgs VEV and   $y^\nu_a$ is the Yukawa coupling between the Higgs and the neutrino. The smallness of neutrino masses can only be pinned on a tiny Yukawa coupling which is not justified by the theory.\\
In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a massive fermion \cite{Majorana1937}. The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$, thus the theory is applicable only to neutral fermions.  Neutrinos are the only neutral elementary particles in the SM -- the only possible Majorana particle candidate. This theory constructs a  neutrino Majorana mass term $\mathcal{L}_5 $ of the following form in the Higgs unitary gauge
\begin{equation}
\mathcal{L}_5 = \frac{1}{2}\frac{gv^2}{\mathcal{M}}\nu^T_L\mathcal{C^\dagger}\nu_L,
 \end{equation}
 where $g$ is the coupling coefficient, $v$ the Higgs VEV and $\mathcal{M}$ a constant with the dimension of the mass proportional to the scale of new physics. The largrangian term $\mathcal{L}_5$ would introduce nonrenormalizable term, since it has dimensions of energy to the fifth power.
 %%%%%%%%%%%%%%%%%
 This is not allowed in the SM lagrangian; however, if the SM is seen as an effective manifestation at low energy 
 an effective theory at low energy 
manifestation of symmetry breaking of a unified theory 
Se si vede il MS come una teoria efficace in una scala di basse energie, manifestazione della rottura di simmetria di una teoria unificata ad alte energie, e non come una teoria definitiva, diventa plausibile l?esistenza di termini non rinormalizzabili nella lagrangiana effettiva; tali termini devono comunque rispettare le simmetrie del MS perché la teoria ad alta energia da cui sono generati deve includere le simmetrie di gauge del MS, in quanto essa si riduce a basse energie al MS stesso.
%%%%%%%%%%%

 In this context, the neutrino mass is given by $m = \frac{g v^2}{\mathcal{M}}$: the coupling coefficient can be of the order of any other coupling coefficient in the SM, since a big value of the new physics mass scale implies small neutrino masses. This vanilla formulation is the conceptual basis for many flavors of theories based on the \emph{see-saw mechanism}, which we will not discuss here in any detail. However, it is fascinating how the puzzle of the neutrino mass hints to the idea of considering the SM as an effective theory at low energy, obtained as the limit of a more complete theory.\\
From a kinematic point of view, Dirac and Majorana neutrinos satisfy the same energy-momentum dispersion relationship. Thus, it is impossible to discern the neutrino nature through kinematic effects such as neutrino oscillations. Neutrinoless double beta decay searches are the most promising way to discern the nature of the neutrino and are therefore subject of great theoretical and experimental interest. Observation of the lepton number violating process  $0\nu\beta\beta$  would imply neutrinos have a Majorana component. Depending on the mass hierarchy, the theory also predicts  $0\nu\beta\beta$  exclusion regions and confirmation of the sole Dirac component for neutrinos\textcolor{red}{find CIT}.\\

>>>>

\subsection{Make up of Neutrino Interactions}
Any neutrino experiment involving the detection of neutrinos is concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
Given the invisible nature of neutrinos, characterizing the products of their interaction is the only method to retrieve their flavor and eventually energy. In real life experiments, this is complicated

@@

\textbf{Majorana or Dirac?}
Evidence of neutrino oscillations demand the introduction in the theory of a mechanism which can give mass to the neutrinos.  This mechanism should possibly also explain why neutrino masses are at least six orders of magnitude lower than the electron mass (the second lightest SM fermion).
In a description of neutrinos as Dirac 4-component spinors, the neutrino field acquires mass  via the Higgs mechanism as any other fermion of the SM. In this case, the neutrino mass is given by 
%\begin{equation} 
$m_a = \frac{y^\nu_a v}{\sqrt2}$, %\end{equation} 
where $v$ is the Higgs VEV and   $y^\nu_a$ is the Yukawa coupling between the Higgs and the neutrino. The smallness of neutrino masses can only be pinned on a tiny Yukawa coupling which is not justified by the theory.\\
In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a massive fermion \cite{Majorana1937}. The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true in the case of two components spinor under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$, thus the theory is applicable only to neutral fermions.  Neutrinos are the only neutral elementary particles in the SM -- the only possible Majorana particle candidate. This theory constructs a  neutrino Majorana mass term $\mathcal{L}_5 $ of the following form in the Higgs unitary gauge
\begin{equation}
\mathcal{L}_5 = \frac{1}{2}\frac{gv^2}{\mathcal{M}}\nu^T_L\mathcal{C^\dagger}\nu_L,
\label{eq:majoMass}
 \end{equation}
 where $g$ is the coupling coefficient, $v$ the Higgs VEV and $\mathcal{M}$ a constant with the dimension of the mass proportional to the scale of new physics. The  $\mathcal{L}_5$ term would introduce a non-renormalizable term in the lagrangian, since it has dimensions of energy to the fifth power.
This is not allowed in the SM term; however, the existence of such terms is plausible if the we consider the SM an effective theory at low energy, manifestation of the symmetry breaking of a Grand Unified Theory (GUT) at higher energy, and not the definitive theory.
The mass term in eq \ref{eq:majoMass} implies the neutrino mass to be  $m = \frac{g v^2}{\mathcal{M}}$. The coupling coefficient can be of the order of any other fermion's coupling coefficient, since the smallness of neutrino masses is achieved by the big value of the new physics mass scale alone. This vanilla formulation is the conceptual basis for many flavors of GUT-based \emph{seesaw mechanism} ~\cite{Yanagida1980}, which we will not discuss here in any detail. However, it is fascinating how the puzzle of the neutrino mass hints to the existence of a deeper and more complete theory.\\
From a kinematic point of view, Dirac and Majorana neutrinos satisfy the same energy-momentum dispersion relationship. Thus, it is impossible to discern the neutrino nature through kinematic effects such as neutrino oscillations. Neutrinoless double beta decay searches are the most promising way to understand the nature of the neutrino and are therefore subject of great theoretical and experimental interest. Observation of the lepton number violating process  $0\nu\beta\beta$  would imply neutrinos have a Majorana component. Depending on the mass hierarchy, the theory also predicts  $0\nu\beta\beta$  exclusion regions and confirmation of the sole Dirac component for neutrinos\textcolor{red}{find CIT}.\\
